Size, Shape and Functionality Dependence of Solute Transport Mechanisms in a Cross-linked H\textsubscript{II} Phase Lyotropic Liquid Crystal Membrane

### 2019-03-05 Ben_Manuscripts/transport/Outline.tex
fixing paths

  \noindent We measured the time-averaged mean squared displacement (MSD)
  of the centers of mass of each solute over the course of 1 $\mu$s MD 
  simulations using Equation~\ref{eqn:tamsd}.
  \begin{itemize}
    \item \begin{equation}
	\overline{x^2(\tau)} = \dfrac{1}{T - \tau}\int_{0}^{T - \tau} (x(t + \tau) - x(t))^2 dt
	\label{eqn:tamsd}
	\end{equation}
	where $\tau$ is the time lag and T is the length of the
	trajectory~\cite{meroz_toolbox_2015}. 
	\item Generally, the MSD grows according to Equation~\ref{eqn:msd_form},
	\begin{equation} 
	\langle x^2(t) \rangle = K_{\alpha}t^\alpha
	\label{eqn:msd_form}
	\end{equation} 
	\item where $\alpha$ is the anomalous exponent and $K_\alpha$ is the
	generalized diffusion coefficient.
	\item A value of $\alpha < 1$ indicates a subdiffusive process, while a value of
	$\alpha = 1$ and $\alpha > 1$ is characteristic of Brownian and superdiffusive
	motion respectively.
	\item In practice, $\alpha$ corresponds to the growth of the \textit{ensemble}
	MSD given by Equation~\ref{eqn:ensemble_msd}~\cite{meroz_toolbox_2015}:
	\begin{equation}
	\langle x^2(t) \rangle = \langle x(t) - x(0) \rangle
	\label{eqn:ensemble_msd}
	\end{equation}
	\item The ensemble MSD is calculated with respect to a reference position and
	hence carries some dependence on its starting point.
	\item The time-averaged MSD averages over all possible time lags of a given 
	length, effectively eliminating any initial configuration dependence and generating
	an increased number of observations. 
	\item For ergodic systems, both types of MSDs will be equal.
	\item Since we have a small number of solutes with which to generate statistics
	and because we are not calculating values for $\alpha$ in this particular study,
	we will only use the time-averaged MSD.
  \end{itemize}

@@

@@

  \subsection*{Structure and Transport of Membrane Components}\label{section:membrane_components}

@@

  In contrast to our previous work with a solvent-free version of this membrane, the pore
  region of the 5 and 10 wt \% water systems are primarily filled with water and sodium
  ions.
  \begin{itemize}
    \item In dry systems, the pore center is densely filled with sodium ions and 
    head groups (See Ref~\cite{coscia_understanding_2019}).
    \item Figure~\ref{fig:component_densities} plots the densities of each
    membrane component.
    \item Water is densest at the center of each pore. 
    \item Sodium ions are densest within the pore region, which we define as the
    space between the pore center and the peak head group density.
    \item The peak density of sodium ions is not at the pore center in either case
    likely because they are still loosely associated with the monomer's 
    carboxylate head groups.
  \end{itemize}

@@

  The MSD of sodium and water become linear on long timescales.
  \begin{itemize}
    \item
  \end{itemize}

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{water_msd_comparison.pdf}
  \caption{Water}\label{fig:water_msd_comparison}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{na_msd_comparison.pdf}
  \caption{Sodium Ions}\label{fig:na_msd_comparison}
  \end{subfigure}
  \caption{}\label{fig:msd_comparison}
  \end{figure}

  Sodium coordinates with water.

  \noindent The diffusion constant of water is about 50 times faster than sodium.
  \begin{itemize}
    \item In a dilute sodium chloride solution, the diffusivity of sodium
    is about 4 times slower than water.
    \item The carboxylate groups, whose oxygen atoms each have a charge of -0.822
    keep the positively charged sodium ions relatively close and restricts their
    motion.
  \end{itemize}

@@

  \noindent The MSDs for the solutes studied in this work span a moderate range. 
  \begin{itemize}
    \item We plotted the 400 ns time-lag MSD for each solute in Figure~\ref{fig:all_msds}.
  	\item The fastest solute, methanol, has an MSD over 20 times greater than
  	the slowest, dimethyl formamide. 
  \end{itemize}

  \noindent The MSDs are not a monotonic function of solute size.
  \begin{itemize}
  	\item We plotted the solute size against their MSDs in Figure~\ref{fig:all_msds}.
  	\item Tetrose, our fifth largest solute, has an average MSD higher than more 
  	than half of all solutes studied. 
  	\item The two slowest solutes, DMF and THF, are smaller than 10 faster solutes.
  	\item Transport is clearly affected by factors other than molecular size. 
  \end{itemize}

  \begin{figure}
  \centering
  \begin{subfigure}{0.5\textwidth}
  \includegraphics[width=\textwidth]{all_10wt_tamsds.pdf}  % include bar plot of MW to show it's non-monotonic
  \caption{}\label{fig:all_msds}
  \end{subfigure}
  \begin{subfigure}{0.5\textwidth}
  \includegraphics[width=\textwidth]{msd_radius.pdf} 
  \caption{}\label{fig:msd_radius}
  \end{subfigure}
  \caption{}\label{fig:msds}
  \end{figure}

@@

  \noindent On the timescales simulated in our study, we observe subdiffusive behavior.
  %, providing the beginnings of an answer to question~\ref{question:mechanisms}.
  \begin{itemize}  
    \item Figure~\ref{fig:example_ztraces} plots the $z$-coordinate versus time of
  	3 representative ethanol centers of mass.
  	\item There are clear periods of entrapment separated by relatively large hops.
	\item The MSD calculated based on all ethanol molecules is plotted in Figure~\ref{fig:example_msd}
	and is sublinear.
	\item The long periods of entrapment likely lead to this sublinear, and thus
	subdiffusive, behavior.
  \end{itemize}

  Solutes partition out of the pore into the head group region and beyond which
  may lead to radially dependent transport mechanisms.
  \begin{itemize}
    \item It is clear from Figure~\ref{fig:example_ztraces} that the longest 
    periods of entrapment generally occur when solutes are far from the pore
    center.
    \item There is a high resistance to movement in the dense head group and
    tail regions.
    \item When hops occur, and where there is the most $z$-positional noise, 
    solutes are generally close to the pore center.
    %BJC: need a radial density of head groups, tails, water etc.
    \item Solutes can move relatively freely when they enter the pore region
    which is primarily composed of water molecules.
  \end{itemize}

@@

  In addition to solute trapping in dense monomer regions, we observe a second
  trapping mechanism caused by preferential hydrogen bonding between 
  hydrogen bond donor solutes and monomer carboxylate groups. 
  \begin{itemize}
    \item Of the time spent donating hydrogen atom(s), x \% of the hydrogen
    bonds occur between solutes and carboxylate groups as opposed to solutes
    with water. 
    \item This shows a high perference for the carboxylate anions since there
    are 5 times more water molecules in the system.
    \item The carboxylate oxygen atoms have a slightly higher partial negative
    charge than water molecules.
    \item Additionally, on average, each sodium ion is coordinated to 3 water
    molecules. Roughly all of the water in the pore region is usually coordinated
    to a sodium ion which decreases their availability to solutes.
  \end{itemize}

@@

>>>>

  \noindent We measured the time-averaged $z$-direction mean squared 
  displacement (MSD) of the centers of mass of each solute over the 
  course of 1 $\mu$s MD simulations using Equation~\ref{eqn:tamsd}.
  \begin{itemize}
    \item \begin{equation}
	\overline{x^2(\tau)} = \dfrac{1}{T - \tau}\int_{0}^{T - \tau} (x(t + \tau) - x(t))^2 dt
	\label{eqn:tamsd}
	\end{equation}
	where $\tau$ is the time lag and T is the length of the
	trajectory~\cite{meroz_toolbox_2015}. 
	\item Generally, the MSD grows according to Equation~\ref{eqn:msd_form},
	\begin{equation} 
	\langle x^2(t) \rangle = K_{\alpha}t^\alpha
	\label{eqn:msd_form}
	\end{equation} 
	\item where $\alpha$ is the anomalous exponent and $K_\alpha$ is the
	generalized diffusion coefficient.
	\item A value of $\alpha < 1$ indicates a subdiffusive process, while a value of
	$\alpha = 1$ and $\alpha > 1$ is characteristic of Brownian and superdiffusive
	motion respectively.
	\item In practice, $\alpha$ corresponds to the growth of the \textit{ensemble}
	MSD given by Equation~\ref{eqn:ensemble_msd}~\cite{meroz_toolbox_2015}:
	\begin{equation}
	\langle x^2(t) \rangle = \langle x(t) - x(0) \rangle
	\label{eqn:ensemble_msd}
	\end{equation}
	\item The ensemble MSD is calculated with respect to a reference position and
	hence carries some dependence on its starting point.
	\item The time-averaged MSD averages over all possible time lags of a given 
	length, effectively eliminating any initial configuration dependence and generating
	an increased number of observations. 
	\item For ergodic systems, both types of MSDs will be equal.
	\item Since we have a small number of solutes with which to generate statistics
	and because we are not calculating values for $\alpha$ in this particular study,
	we will only use the time-averaged MSD.
  \end{itemize}

@@

  Using four different methods, Rowley and Roux observed peaks in the radial
  distribution function for sodium coordinated with water at an O-Na distance 
  of between 2.3 and 2.5 \AA~\cite{rowley_solvation_2012}. We used 2.5 \AA as
  the distance cutoff.

@@

  \subsection*{Structure and Transport of Membrane Constituents}\label{section:membrane_components}

@@

  In contrast to our previous work with a solvent-free version of this membrane, the pore
  region of the 5 and 10 wt \% water systems are primarily filled with water and sodium
  ions.
  \begin{itemize}
    \item In dry systems, the pore center is densely filled with sodium ions and 
    head groups (See Ref~\cite{coscia_understanding_2019}).
    \item Figure~\ref{fig:component_densities} plots the densities of each
    membrane constituent.
    \item Water is densest at the center of each pore. 
    \item Sodium ions are densest between the pore center and the peak head group density.
    \item The peak density of sodium ions is not at the pore center in either case
    likely because they are still loosely associated with the monomer's 
    carboxylate head groups.
  \end{itemize}

@@

  The MSD of sodium and water is significantly larger in the 10 wt \% water system
  versus the 5 wt \% water system.
  \begin{itemize}
    %BJC: these numbers will change. This is for incomplete methanol systems. I will
    % redo calculations on pure water systems.
    \item Water moves about 65 times faster and sodium moves about 40 times faster
    in the 10 wt\% system.
    \item Water moves about 65 times faster than sodium in the 10 wt \% system and
    40 times faster than sodium in the 5 wt \% system. %BJC: weird coincidence
    \item The diffusivity of water in the 10 wt \% system is still only 1 \% that of 
    bulk water %BJC: 5.53e-07 versus 5.5e-05 cm^2/s
  \end{itemize}

  % BJC: could talk about subdiffusion a little here but I'd rather save it for 
  % small molecule discussion

%  \noindent The diffusion constant of water is about 50 times faster than sodium.
%  \begin{itemize}
%    \item In a dilute sodium chloride solution, the diffusivity of sodium
%    is about 4 times slower than water.
%    \item The carboxylate groups, whose oxygen atoms each have a charge of -0.822
%    keep the positively charged sodium ions relatively close and restricts their
%    motion.
%  \end{itemize}

  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{water_msd_comparison.pdf}
  \caption{Water}\label{fig:water_msd_comparison}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{na_msd_comparison.pdf}
  \caption{Sodium Ions}\label{fig:na_msd_comparison}
  \end{subfigure}
  \caption{(a) The MSD of water in the 10 wt \% water system is about 65 times
  higher than water in the 5 wt \% water system. (b) The MSD of sodium in the
  10 wt \% water system is about 40 times higher than sodium in the 5 wt \% 
  water system.}\label{fig:msd_comparison}
  \end{figure}

  \noindent Sodium coordinates with water far less than in bulk solution.
  \begin{itemize}
    \item On average, each sodium ion is coordinated to about 1.7 water molecules in 
    the 10 wt \% system and 1.15 water molecules in the 5 wt\% system.
    \item In bulk water, it is estimated that sodium coordinates with 5.6 - 5.8 water
    molecules~\cite{rowley_solvation_2012}
    \item The low levels of water coordination under confinement are likely driven
    in part by favorable interactions between sodium ions and carboxylate head groups.
    \item A dehydrated sodium ion is not shielded from the negatively charged carboxylate
    group.~\cite{ma_drastically_2019}
  \end{itemize}

  % BJC: This figure is from methanol simulations. Will be replaced by pure water
  % simulations
  % BJC: I can probably just report the averages rather than plot this
  \begin{figure}[!htb]
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{Na_water_coordination.pdf}
  \caption{}\label{fig:na_water_coordination}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{Na_hg_coordination.pdf}
  \caption{}\label{fig:na_hg_coordination}
  \end{subfigure}
  \end{figure}

@@

  We observe trends in transport properties that are dependent on chemical 
  environment within the nanopores as well as the chemical functionality of membrane
  constituents rather than just solute size. 
  \begin{itemize}
    \item Polar solutes are slowed by interactions between monomer functional groups
    and ions.
    \item Solutes with hydrophobic character partition out of the pore and are 
    slowed by densely packed organic monomer components.
    \item A thorough understanding of these interactions will help us to create
    monomer design principles.
    \item We will begin our analysis by considering the collective trends observed  
    across all systems and then focus on subsets of similar molecules where 
    some exceptions occur. 
  \end{itemize}

@@

  Like water and sodium above, the MSDs of the solutes studied in this work are 
  significantly larger in the 10 wt \% system, than those in the 5 wt \% water 
  system (Figure~\ref{fig:msds}). 
  \begin{itemize}
    \item The fastest moving solute in both cases, methanol, has an MSD about 175
    times larger in the 10 wt \% versus the 5 wt \% system.
    \item Clearly the equilibrium water content of a given LLC system will 
    determine its viability for real separations.
  \end{itemize}

%  \noindent The MSDs for the solutes studied in this work span a moderate range. 
%  \begin{itemize}
%    \item We plotted the 400 ns time-lag MSD for each solute in Figure~\ref{fig:all_msds}.
%  	\item The fastest solute, methanol, has an MSD over 20 times greater than
%  	the slowest, dimethyl formamide. 
%  \end{itemize}

  \noindent The MSDs are not a monotonic function of solute size.
  \begin{itemize}
  	\item We plotted the solute size against their MSDs in Figure~\ref{fig:msd_radius_5wt} and~\ref{fig:msd_radius_10wt}.
  	\item There are numerous examples of solutes in our study that do not 
  	follow the behavior set forth by the Stokes-Einstein equation which 
  	implies that solute MSD should be inversely related to solute size.  % citation
  	\item This behavior is not entirely surprising but worth emphasizing. 
  	\item It is expected that the Stokes-Einstein relationship, which assumes that 
  	the solute radius is much larger than the solvent, should break down once
  	solute and solvent are of comparable size.
  	\item Membranes, including that studied here, are often characterized 
  	with size-exclusion experiments which emphasizes selectivity based on the ability
  	to enter a given pore. 
  	\item Once we reach the nm to sub-nm pore size regime, size-exclusion 
  	becomes of secondary importance since the primary purpose of these membranes is to 
  	separate molecules whose sizes are on the same order as water.
  	\item In order to handle complex waste streams with molecules of similar size,
  	it is important to take advantage of material properties other than pore size. 
%  	\item Tetrose, our fifth largest solute, has an average MSD higher than more 
%  	than half of all solutes studied. 
%  	\item The two slowest solutes, DMF and THF, are smaller than 10 faster solutes.
%  	\item Transport is clearly affected by factors other than molecular size. 
  \end{itemize}

  \begin{figure}
  \centering
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{all_5wt_tamsds.pdf}
  \caption{5 wt \% water}\label{fig:all_msds_5wt}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{all_10wt_tamsds.pdf}
  \caption{10 wt \% water}\label{fig:all_msds_10wt}
  \end{subfigure}
  %BJC: missing a few since those simulations just got started
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_radius_5wt.pdf} 
  \caption{5 wt \% water}\label{fig:msd_radius_5wt}
  \end{subfigure}
  \begin{subfigure}{0.45\textwidth}
  \includegraphics[width=\textwidth]{msd_radius_10wt.pdf} 
  \caption{10 wt \% water}\label{fig:msd_radius_10wt}
  \end{subfigure}
  \caption{}\label{fig:msds}
  \end{figure}

  \noindent On the timescales simulated in our study, we observe subdiffusive behavior.
  %, providing the beginnings of an answer to question~\ref{question:mechanisms}.
  \begin{itemize}  
    \item Figure~\ref{fig:example_ztraces} plots the $z$-coordinate versus time of
  	3 representative ethanol centers of mass in the 10 wt \% water system.
  	\item There are clear periods of entrapment separated by relatively large hops.
	\item The MSD calculated based on all ethanol molecules is plotted in 
	Figure~\ref{fig:example_msd} and is sublinear.
	\item The long periods of entrapment likely lead to this sublinear, and thus
	subdiffusive, behavior.
  \end{itemize}

  Solutes partition out of the pore into the head group region and beyond which
  may lead to radially dependent transport mechanisms.
  \begin{itemize}
    \item It is clear from Figure~\ref{fig:example_ztraces} that the longest 
    periods of entrapment usually occur when solutes are far from the pore
    center.
    \item There is a high resistance to movement in the dense head group and
    tail regions.
    \item When hops occur, and where there is the most $z$-positional noise, 
    solutes are generally close to the pore center.
    \item Solutes can move relatively freely when they enter the pore region
    which is primarily composed of water molecules and sodium ions.
  \end{itemize}

@@

  %BJC: need to check whether sodium ion association is exhibited by all solutes
  In addition to solute trapping in dense monomer regions, we observe a second
  trapping mechanism caused by preferential hydrogen bonding between 
  hydrogen bond donor solutes and monomer carboxylate groups. 
  \begin{itemize}
    \item Of the time spent donating hydrogen atom(s), x \% of the hydrogen
    bonds occur between solutes and carboxylate groups as opposed to solutes
    with water. 
    \item This shows a high preference for the carboxylate anions since there
    are 5 times more water molecules in the system.
    \item The carboxylate oxygen atoms have a slightly higher partial negative
    charge than water molecules.
    \item Additionally, on average, each sodium ion is coordinated to 3 water
    molecules. Roughly all of the water in the pore region is usually coordinated
    to a sodium ion which decreases their availability to solutes.
  \end{itemize}

@@

  \noindent The transport behavior exhibited by solutes in the 5 wt \% water systems
  is similar to those shown by those in the 10 wt \% system however the timescales are
  much longer.
  \begin{itemize}
    %BJC: will put supporting data in supporting information
    \item We observe subdiffusive behavior with intermittent hopping between
    periods of entrapment.
    \item The frequency and length of hops are both diminished in the 5 wt \% system.
    \item Since there are only 24 solute molecules in each system, in order to obtain
    better time-averaged descriptions of solute transport mechanisms, we will focus
    the rest of our analysis on transport in the 10 wt \% water systems.
  \end{itemize}
